### 2013-07-12 mcmc.tex
convergence

The reason that there is no simple answer is that you can't really
ever know that you have sampled the full posterior pdf, and the reason
for \emph{that} is that if you \emph{could} know that, you would also
be able to solve the famously difficult discrete optimization
problem.\note{explain this remark and cite something here}
Another way to put it is that you might have two modes---two separated
regions of substantial total probability---in the posterior pdf that
are both important but separated by a large region of low probability.
If you are using a simple sampler, you could sample one of these modes
very well but the sampler will take effectively infinite time to find
the other mode.
Any test of convergence would be passed, but in fact the sampling
would be incomplete and not representative.
That is, knowing that you have a complete and representative sampling
is nearly impossible.

Heuristically, however, you have sampled long enough when you can see
that the (or each) walker has traversed the high-probability parts of
the parameter space many times in the length of the chain.
Or, equivalently, you have sampled long enough when the first half of
the chain shows very much the same posterior pdf morphology as the
second half of the chain, or indeed as any substantial subset.

>>>>

The reason that there is no simple answer is that you can't really
ever know that you have sampled the full posterior pdf, and the reason
for \emph{that} is that if you \emph{could} know that, you would also
be able to solve the famously difficult discrete optimization
problem.\note{explain this remark and cite something here}
Another way to put it is that you might have two modes---two separated
regions of substantial total probability---in the posterior pdf that
are both important but separated by a large region of low probability.
If you are using a simple sampler, you could sample one of these modes
very well but the sampler will take effectively infinite time to find
the other mode.
Any test of convergence would be passed, but in fact the sampling
would be incomplete and not representative.
That is, knowing that you have a complete and representative sampling
is effectively impossible.

In this context,
  it is \emph{simply not fair} to require an MCMC run to have fully and completely
  sampled the posterior pdf,
  at least not in any provable sense.
The question of whether a sampling is converged to a representative sampling
  of the posterior pdf is actually outside the domain of science,
  because the domain of science is the domain of questions answerable by scientists.
That is, you can \emph{never know} that you have correctly sampled the posterior pdf,
  despite many statements in the literature to the contrary.%
  \note{Give example statements... Mention cases in which you \emph{might} be able to claim this...}
The upshot of this is that we don't and can't require any kind of absolute convergence;
  indeed, it would be impossible even to test for it.

In this pragmatic situation---the Real World, as it is called---we have to rely on heuristics.
Heuristically, you have sampled long enough when you can see
that the (or each) walker has traversed the high-probability parts of
the parameter space many times in the length of the chain.
Or, equivalently, you have sampled long enough when the first half of
the chain shows very much the same posterior pdf morphology as the
second half of the chain, or indeed as any substantial subset.
